Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset

### 2024-09-24 code/manuscript.Rmd
Updated manuscript for running cont savings

output:
  bookdown::word_document2:
    reference_docx: "../paper/TF_Template_Word_Mac_2011.dotx"
editor_options: 
  markdown: 
    wrap: 72
# bibliography: references.bib
# biblio-style: apalike

@@

df_energy_tidy <- read_rds(paste0(readfile_tidy, "df_energy.rds"))
df_meta_tidy <- read_rds(paste0(readfile_tidy, "df_meta.rds"))
df_weather_tidy <- read_rds(paste0(readfile_tidy, "df_weather.rds"))
df_sprt_all_tidy <- read_rds(paste0(readfile_tidy, "df_sprt_all.rds"))
df_seq_FS_tidy <- read_rds(paste0(readfile_tidy, "df_seq_FS.rds"))
df_NRE_occ <- read_rds(paste0(readfile_tidy, "df_NRE_occ.rds"))
df_MD_tidy <- read_rds(paste0(readfile_tidy, "df_MD.rds"))
df_FS_tidy <- read_rds(paste0(readfile_tidy, "df_FS.rds"))
df_eui_tidy <- read_rds(paste0(readfile_tidy, "df_eui.rds"))
df_cont_MD_tidy <- read_rds(paste0(readfile_tidy, "df_cont_MD.rds"))
df_cont_FS_tidy <- read_rds(paste0(readfile_tidy, "df_cont_FS.rds"))

@@

df_energy_messy <- read_rds(paste0(readfile_messy, "df_energy.rds"))
df_meta_messy <- read_rds(paste0(readfile_messy, "df_meta.rds"))
df_weather_messy <- read_rds(paste0(readfile_messy, "df_weather.rds"))
df_sprt_all_messy <- read_rds(paste0(readfile_messy, "df_sprt_all.rds"))
df_seq_FS_messy <- read_rds(paste0(readfile_messy, "df_seq_FS.rds"))
df_MD_messy <- read_rds(paste0(readfile_messy, "df_MD.rds"))
df_FS_messy <- read_rds(paste0(readfile_messy, "df_FS.rds"))
df_eui_messy <- read_rds(paste0(readfile_messy, "df_eui.rds"))
df_cont_MD_messy <- read_rds(paste0(readfile_messy, "df_cont_MD.rds"))
df_cont_FS_messy <- read_rds(paste0(readfile_messy, "df_cont_FS.rds"))
df_FS_nsprt <- read_rds(paste0(readfile_messy, "df_FS_nsprt.rds"))
df_MD_nsprt <- read_rds(paste0(readfile_messy, "df_MD_nsprt.rds"))
df_seq_FS_nsprt <- read_rds(paste0(readfile_messy, "df_seq_FS_nsprt.rds"))

@@

@@

| Parameter              | Description                                                      | Value |
|-----------------|--------------------------------------|-----------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

@@

@@

@@

>>>>

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
# bibliography: references.bib
# biblio-style: apalike

@@

df_energy_tidy <- read_rds(paste0(readfile_tidy, "df_energy.rds"))
df_meta_tidy <- read_rds(paste0(readfile_tidy, "df_meta.rds"))
df_weather_tidy <- read_rds(paste0(readfile_tidy, "df_weather.rds"))
df_sprt_all_tidy <- read_rds(paste0(readfile_tidy, "df_sprt_all.rds"))
df_seq_FS_tidy <- read_rds(paste0(readfile_tidy, "df_seq_FS.rds"))
df_NRE_occ <- read_rds(paste0(readfile_tidy, "df_NRE_occ.rds"))
df_MD_tidy <- read_rds(paste0(readfile_tidy, "df_MD.rds"))
df_FS_tidy <- read_rds(paste0(readfile_tidy, "df_FS.rds"))
df_eui_tidy <- read_rds(paste0(readfile_tidy, "df_eui.rds"))
df_cont_MD_tidy <- read_rds(paste0(readfile_tidy, "df_cont_MD.rds"))
df_cont_FS_tidy <- read_rds(paste0(readfile_tidy, "df_cont_FS.rds"))
df_model_acc_tidy <- read_rds(paste0(readfile_tidy, "df_model_acc.rds"))

@@

df_energy_messy <- read_rds(paste0(readfile_messy, "df_energy.rds"))
df_meta_messy <- read_rds(paste0(readfile_messy, "df_meta.rds"))
df_weather_messy <- read_rds(paste0(readfile_messy, "df_weather.rds"))
df_sprt_all_messy <- read_rds(paste0(readfile_messy, "df_sprt_all.rds"))
df_seq_FS_messy <- read_rds(paste0(readfile_messy, "df_seq_FS.rds"))
df_MD_messy <- read_rds(paste0(readfile_messy, "df_MD.rds"))
df_FS_messy <- read_rds(paste0(readfile_messy, "df_FS.rds"))
df_eui_messy <- read_rds(paste0(readfile_messy, "df_eui.rds"))
df_cont_MD_messy <- read_rds(paste0(readfile_messy, "df_cont_MD.rds"))
df_cont_FS_messy <- read_rds(paste0(readfile_messy, "df_cont_FS.rds"))
df_FS_nsprt <- read_rds(paste0(readfile_messy, "df_FS_nsprt.rds"))
df_MD_nsprt <- read_rds(paste0(readfile_messy, "df_MD_nsprt.rds"))
df_seq_FS_nsprt <- read_rds(paste0(readfile_messy, "df_seq_FS_nsprt.rds"))
df_model_acc_messy <- read_rds(paste0(readfile_messy, "df_model_acc.rds"))

@@

### Measurement and verification

### Randomized experimental design

@@

| Parameter              | Description                                                      | Value |
|------------------|------------------------------------|------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

The outbreak of pandemic in 2020 followed by working-from-home policy is
a practical example of building energy consumption influenced by
non-routine events. During unoccupied months, building energy
consumption dropped compared to the same time period in the previous
years. If an energy-efficient intervention was deployed onsite starting
at 2020, given by the whole-building electricity measurements, the large
difference in measured electricity usage in 2020 and projected baseline
usage in 2019 can hardly indicate the retrofit intervention effect.

In other words, if there is no intervention applied or the intervention
effect is known to be null (e.g. constantly being overriden by the
baseline), an ideal M&V method should detect no savings. To test this,
we kept the randomized schedule as sampled earlier but removed added
chilled water plant intervention effect. Therefore, all sampled
intervention days are considered additional sampled baseline days.
Earlier we mentioned the 'messy' dataset contains some change in the
two-year electricity measurement, which is purely due to various
non-routine events such as occupancy change or even other measures
implemented by the building manager but irrelevant to the chilled water
set point reset intervention. Thus, when the proposed chilled water
reset intervention was removed, a more reliable M&V method should
overcome the influence of those confounding factors and inform no
savings.

```{r nosavings, fig.cap="Distribution of savings estimation results by the convention M&V method and randomized method.", fig.width=8, fig.height=4}

df_FS_nsprt %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  left_join(df_seq_FS_nsprt %>% filter(seq == "eob"), by = c("name", "site")) %>% 
  pivot_longer(c(true, conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
  filter(parameter != "true") %>% 
  mutate(parameter = as.factor(parameter), 
         parameter = recode_factor(parameter, 
                                   "conv" = "Conventional M&V", 
                                   "rand" = "Randomized M&V", 
                                   "FS" = "Randomized M&V at early stop")) %>% 
  ggplot(aes(x = parameter, y = value, fill = parameter)) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 1.2) +
  annotate(geom = "text", 
           x = 0.5, 
           y = 0.5, 
           color = 'red',
           label = "True savings = 0") +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 5), 
                     limits = c(0, 16), 
                     labels = number_format(suffix = "%")) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       title = "Savings estimation comparison", 
       subtitle = "between convention and randomized M&V (with early stop)") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

Figure \@ref(fig:nosavings) shows that the conventional M&V method in
average detected 5% savings when there should be none. The randomized
method even when stops right after satisfying all criteria, the savings
estimation is much closer to 0%. This is mostly because when there is a
change or disturbance applied to the target building such as lighting
retrofit or thermal leakage in the envelope, randomly sample the two
strategies at 50%/50% ensures the resulting effect either as a decrease
in lighting electricity or an increase in reheat electricity is balanced
between the two implemented strategies.

@@

Although this paper argues that regression model prediction results can
be largely biased by various non-routine events, we further clarify that
this is not related to regression model fitting accuracy. Time-of-week
temperature model is used here for its simplicity and convenience since
the model only requires outdoor temperature measurements and time of
week as independent variables. The model assumes a linear composition of
building energy consumption as temperature-dependent and time-dependent
load. The time-dependent component accounts for day-to-day variations
and the temperature dependent component considers a piecewise linear
relationship across a range of temperature intervals.
<!--# TODO: add reference -->

To assess modeling accuracy, we used the Coefficient of Variation of
Root-Mean Squared Error, or CV(RMSE) as the error metric. Since this
metric is calculated as a normalized value, it is useful to compare
different model fitting results. Guideline 14 requires that
whole-building baseline model fitting accuracy should maintain a
CV(RMSE) lower than 30%. In addition One study focusing on the baseline
energy data-driven model fitting indicates that TOWT performs as
accurate as other more advanced machine learning models
<!--# TODO: reference --> and the calculated CV(RMSE) distribution for a
large sample of commercial buildings indicates a median of 20%. Figure
\@ref(fig:towtacc) plots the distribution of the model fitting accuracy
calculated separately for the two measurement sets. The result shows
that the TOWT model performance is even better in this study compared to
the literature and no significant difference between the two subsets.
This is mostly related to low-quality measurement sets filtering in the
pre-processing mentioned in Section \@ref(method).

```{r towtacc, fig.cap="TOWT model fitting accuracy distribution on filtered measurement set", fig.width=8, fig.height=4}

df_model_acc_messy %>% 
  mutate(group = "messy") %>% 
  rbind(df_model_acc_tidy %>% mutate(group = "tidy")) %>% 
  mutate(group = as.factor(group)) %>% 
  ggplot(aes(x = group, y = cvrmse, fill = group)) +
  stat_boxplot(geom ='errorbar', width = 0.5) +
  geom_boxplot(outlier.shape = NA) +
  geom_text(aes(x = group, y = median, label = paste0(round(median, digits = 0), " %")), 
            data = . %>% 
              group_by(group) %>% 
              summarise(median = median(cvrmse)) %>% 
              ungroup(), 
            position = position_nudge(y = 1)) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %")) +
  coord_cartesian(ylim = c(0, 38)) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       title = "TOWT model accuracy", 
       subtitle = "CV(RMSE)") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

In general, the limitation is that despite the regression model tends to
capture well on the mean building energy consumption but can
underestimate and peak or overestimate the lower base load
<!--# TODO: reference -->, which is not useful for assessing demand
response. Additionally, a data-driven model can learn the training
dataset very efficiently with high accuracy, but the well-trained model
is challenging to generalize prediction outside the range of training
set.

@@

Another advantage of using randomized M&V is the flexibility of changing
sampling ratio after the M&V. For example, the building owner can
continue sampling at 50%/50% between the baseline and the intervention
to further reduce the uncertainty range associated with the savings. Or
he/she can switch to 100% intervention to optimize energy savings, but
whether the existing baseline becomes outdated is unknown. A compromised
approach is to continue sampling at 80%/20% between the baseline and the
intervention. Figure \@ref(fig:cont) shows after the analyst reports the
randomized M&V results, a new schedule sampling at 80%/20% was
implemented for another 36 weeks.

```{r cont, fig.cap="Savings estimation accuracy for continue sampling at 80%/20% for 36 weeks after all stopping criteria are satisfied", fig.width=8, fig.height=4}

plot_data <- df_cont_FS_tidy %>% 
  rename(cont = FS) %>% 
  left_join(df_FS_tidy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(abs_diff = abs(savings - cont), 
         plot_max = max(abs_diff))

p1 <- plot_data %>% 
  arrange(abs_diff) %>% 
  ggplot() +
  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff), 
           stat = "identity", alpha = 0.5) +
  geom_hline(yintercept = median_rand, 
             linetype = "dashed", 
             color = "red") +
  annotate(y = median_rand + 0.5, 
           x = 20, 
           geom = "text", 
           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
           size = 4) +
  geom_hline(yintercept = upper_rand, 
             linetype = "dashed", 
             color = "red") +
  annotate(y = upper_rand + 0.5, 
           x = 20, 
           geom = "text", 
           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
           size = 4) +
  scale_x_continuous(expand = c(0.02, 0), 
                     breaks = breaks_pretty(n = 3)) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %"),
                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
  labs(x = "Number of buildings", 
       y = "Absolute difference in measured savings", 
       subtitle = "Tidy subset") +
  coord_cartesian(ylim = c(0, 25)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

plot_data <- df_cont_FS_messy %>% 
  rename(cont = FS) %>% 
  left_join(df_FS_messy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(abs_diff = abs(savings - cont), 
         plot_max = max(abs_diff))

p2 <- plot_data %>% 
  arrange(abs_diff) %>% 
  ggplot() +
  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff), 
           stat = "identity", alpha = 0.5) +
  geom_hline(yintercept = median_rand, 
             linetype = "dashed", 
             color = "red") +
  annotate(y = median_rand + 0.5, 
           x = 220, 
           geom = "text", 
           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
           size = 4) +
  geom_hline(yintercept = upper_rand, 
             linetype = "dashed", 
             color = "red") +
  annotate(y = upper_rand + 0.5, 
           x = 220, 
           geom = "text", 
           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
           size = 4) +
  scale_x_continuous(expand = c(0.02, 0), 
                     breaks = breaks_pretty(n = 3)) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %"),
                     limits = c(0, max(plot_data$plot_max) + 0.1)) +
  labs(x = "Number of buildings", 
       y = NULL, 
       subtitle = "Messy subset") +
  coord_cartesian(ylim = c(0, 25)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        axis.text.y = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 0, unit = "mm"))

ggarrange(p1, p2,
          ncol=2, nrow=1,
          labels = c("a)", "b)"),
          align = "hv",
          legend="bottom") +
  plot_annotation(title = str_glue("Continue sampling accuracy comparison"), 
                  subtitle = "with 80%/20% intervention/baseline")
```

The deviation remains consistent in the tidy subset but increases
slightly in the messy subset, suggesting that changes in energy usage in
the target building create a trade-off between measured savings and
estimation accuracy when down-sampling the baseline. In this scenario,
opting for an early stop at 24 or 36 weeks with 50% intervention and
resampling 80% of the data for 36 weeks still allows the building owner
to capture 65% of the full-range savings.

@@

Considering the diversity of the measurement sets, the assumption made
in this study is relative general. For example,
